---
title: "Homework 1"
subtitle: "IQM"
date: "`r strftime(Sys.Date(), format = '%Y-%m-%d')`"
author: Patrick Anker
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{units}
- \usepackage{tikz}
- \usepackage{bm}
- \usepackage{graphicx}
- \usepackage{subcaption}
output: pdf_document
---

```{r setup, include=FALSE, cache=FALSE}
knitr::read_chunk(here::here("homework1/analysis.R"))
set.seed(0xABBA)
```

# Question 1

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.4\linewidth}
    \includegraphics[width=\linewidth]{`r here::here("homework1/q1_sig3-9.jpg")`}
    \caption{$\sigma \approx 3.9$}
  \end{subfigure}
  \begin{subfigure}[b]{0.4\linewidth}
    \includegraphics[width=\linewidth]{`r here::here("homework1/q1_sig10.jpg")`}
    \caption{$\sigma \approx 10$}
  \end{subfigure}
  \caption{Line sketches of $y=10x + 30$ with differing standard deviations}
\end{figure}

# Question 2

### (a) Forecasting/classification

Everyday, millions of New Yorkers travel by the subway system. Using some metric of turnstile entries, the MTA could forecast the number of trains needed to be scheduled based on the passenger demand as a basic start. It could further model required trains by tracking average attendance to larger events across the city, e.g. Yankees or Nets games.

### (b) Exploring associations

As the climate crisis intensifies, modes of transportation and city planning need drastic changes to rapidly reduce the need and demand for automobiles (internal combustion engine automobiles at the very least). Advocates commonly propose e-bikes as an alternative to cars; however, necessary infrastructure needs to be put in place to make that transition happen. Often, car users come from further away where transit and bike access are limited. The NYC Department of Planning could examine the relationship of transit access and bike ridership to see if they are related. Having this knowledge could inform planning decisions for resource allocation, e.g. just putting in a bike lane or reconfiguring a major right of way to be bus-only while adding bike lanes in parallel.

### (c) Extrapolation

Following from **(b)**, the city also needs to make policy decisions that encourage the use of alternative modes like e-bikes. A common roadblock to uptake of differing modes is the lack of financial incentive. A survey could be implemented to examine the addition of a possible tax credit for purchasing one's own e-bike versus adding additional funds to the CitiBike program for its e-bike transition. The results of this survey, if appropriately sampled, could be extrapolated to the general NYC populace and help the City Council make a targeted policy shift based on what New Yorkers would actually prefer.

### (d) Causal Inference

The classic traffic engineering problem is **induced demand**: the more there is infrastructure for a mode, the more likely that mode will be used -- inducing demand. This has ensnared American urban design for the last 80 years and needs to be addressed head-on. Locally, an experiment could be conducted: if a roadway is closed to vehicular traffic, do neighboring transit or micromobility (e.g. bikeshare) modes see an increase in use? Since the Covid pandemic, we have seen an explosion in popularity of these kinds of "Quiet Streets" treatments. This causal question is vital to be understood for planners to create low-cost programs across the city to reverse the induced demand problem from cars to human-scale transport.

# Question 3

### Does air pollution cause lung damage?

In this question, we're examining the effect of air pollution on the lungs. The predictor would be some measure of air pollution (e.g. levels of PM2.5 particulates), and the outcome would be a measure of lung tissue damage (perhaps repeated MRIs and peak air flow tests). If we were to use MRIs, I would expect to see a trend where as average PM2.5 counts increase, the amount of tissue occlusion (signifying damage) would also increase; if we were to use peak air flow tests, I would expect to see a trend where as average PM2.5 counts increase, peak air flow values would decrease. For the peak airflow measure, we would have to compensate for altitude.

### Is it true that people who exercise more are healthier?

The predictor in this situation would be something like average time spent exercising per week. We could be more precise and include how much time the heart rate is an exercise zone, but for longer term effects on health, the average time spent exercising is probably good enough. The outcome variable is more subtle: we could use something like some BMI measure, resting heart rate, or VO2 max. Each on their own is not a direct measure of health but they can point to a healthier lifestyle. For BMI and resting heart rate, I would expect to see trend comparing average time spent exercising to either BMI (assuming the baseline measurement was on the heavier side) or resting heart rate to be downward.

### Are grapes sweeter the longer they sit on the vine?

The predictor for this situation is pretty straightforward: the length of time a grape was on the vine. However, the outcome is more tricky. A naive approach would be to examine sugar content in the picked grapes, but the term "sweeter" is a subjective measure of taste, and flavor profiles can be moderated by all sorts of volatile compounds. I would probably administer a blinded taste test with a Likert scale of sweetness as the outcome (lower values being less sweet). With an ordinal outcome, one would have to use transformation like an ordered logit with a cumulative link function; after transforming the data, I would generally expect to see that as the average time on the vine increases, the transformed values would increase (showing that proportionally more people thought the sampled grapes were sweeter).

# Question 4

```{r, include=FALSE}
props <- c(.5, .6, .4, .3)
weights <- c(200, 250, 300, 250)

avg <- round(weighted.mean(props, weights), 2)
```

Using the weighted averages formula

$$
E[x] = \frac{\sum_i x_i w_i}{\sum_i w_i}
$$

we can substitute in with the proportions per each cohort:

$$
E[p] = \frac{.5*200 + .6*250 + .4*300 + .3*250}{200 + 250 + 300 + 250}
$$

which evaluates to \(E[p] = \) `r avg`.

# Question 5

Since we are free to choose the weights, we can oversample the cohorts that are not in favor of higher taxes. While we do not know the actual proportions of each cohort in the city's population, we can aim for the 40% figure as a guide.

We only have one equation, so effectively three of the weights can been arbitrarily chosen. For simplicity, let's choose $w_{\mathrm{18-29}} = 1$, $w_{\mathrm{30-44}} = 1$, and $w_{\mathrm{45-64}} = 2$ (because we want to oversample those who don't like raising taxes).

This leaves the weight for the 65+ cohort. Some simple algebra will show that value:

\begin{align*}
.4\ &=\ \frac{.5*1 + .6*1 + .4*2 + .3*w}{4 + w} \\
&=\ \frac{.3w + 1.9}{4 + w} \\
w+4 &=\ \frac{.3w + 1.9}{.4} \\
w\left(1 - \frac{3}{4}\right)\ &=\ 4.75 - 4 \\
\frac{w}{4}\ &=\ \frac{3}{4} \\
w\ &=\ 3
\end{align*}

For practicality, let's multiply the weights by 200; therefore our target cohort sizes are:

```{r, echo=FALSE}
tidytable::tidytable(
  Cohort = c("18-29", "30-44", "45-64", "65+"),
  "Weight (target recruitment size)" = c(200, 200, 400, 600)
) |>
  knitr::kable()
```

# Question 6

```{r question6, echo=FALSE}
```

# Question 7

We can approximate the distribution of grades to be roughly normal given that the process of adding points on an exam is roughly a random walk. Linear transformations on normal distributions are normal. Therefore, algebraically, we can compare unitless z-scores to find the linear transformation.

\begin{align*}
\frac{y-\mu^\prime}{\sigma^\prime}\ &=\ \frac{x-\mu}{\sigma} \\
\sigma y\ &=\ \sigma^\prime x - \sigma^\prime \mu + \sigma\mu^\prime \\
y\ &=\ \frac{\sigma^\prime}{\sigma}x + \left(\mu^\prime - \frac{\sigma^\prime}{\sigma}\mu\right)
\end{align*}

Substituting $\mu = 35$, $\mu^\prime = 100$, $\sigma = 10$, and $\sigma^\prime = 15$,

\begin{align*}
y(x)\ &= 1.5x + 47.5\ \rightarrow\ \{a = 47.5,\ b = 1.5\} &\ \textbf{(a)}
\end{align*}

To find the range, we need to insert the limits of $x$, 0 and 50:

\begin{align*}
y(0)\ &=\ 1.5(0) + 47.5\ \rightarrow y_{\mathrm{min}}=47.5 \\
y(50)\ &=\ 1.5(50) + 47.5\ \rightarrow y_{\mathrm{max}}=122.5 \\
\end{align*}

Thus, the range of $y$ is $\left[47.5, 122.5\right]$. **(b)**

A plot of the linear transformation follows: **(c)**

```{r question7, echo=FALSE, fig.height=4}
```

# Question 8

<!-- For the definitions -->
```{r question8a, include=FALSE}
```

Using the `dat_bass` data, we can compute the mean and standard deviation of bass lengths (assumed to be millimeters) by age:

```{r, echo=FALSE}
dat_bass |>
  tidytable::select(
    "age (years)" = Age,
    "mean length (mm)" = length_mean,
    "std. dev. length (mm)" = length_sd,
  ) |>
  unique() |>
  tidytable::arrange(`age (years)`) |>
  knitr::kable()
```

As shown in the table, there appears to be a somewhat linear trend between age and standard deviation of length. The following plot focuses on that relationship:

```{r question8b, echo=FALSE}
```

Presumably, the uncertainty increases over time because fewer and fewer fish are available to be measured, and according to the table, it seems the fish that live longer are also physically larger. It follows that perhaps the fish that lived long enough to be sampled were also some of the healthier fish.

