---
title: "Homework 1"
subtitle: "IQM"
author: Patrick Anker
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{units}
- \usepackage{tikz}
- \usepackage{bm}
output: pdf_document
---

```{r setup, include=FALSE, cache=FALSE}
knitr::read_chunk(here::here("homework1/analysis.R"))
set.seed(0xABBA)
```

# Question 1

# Question 2

# Question 3

# Question 4

# Question 5

# Question 6

```{r question6, echo=FALSE}
```

# Question 7

We can approximate the distribution of grades to be roughly normal given that the process of adding points on an exam is roughly a random walk. Linear transformations on normal distributions are normal. Therefore, algebraically, we can compare unitless z-scores to find the linear transformation.

\begin{align*}
\frac{y-\mu^\prime}{\sigma^\prime}\ &=\ \frac{x-\mu}{\sigma} \\
\sigma y\ &=\ \sigma^\prime x - \sigma^\prime \mu + \sigma\mu^\prime \\
y\ &=\ \frac{\sigma^\prime}{\sigma}x + \left(\mu^\prime - \frac{\sigma^\prime}{\sigma}\mu\right)
\end{align*}

Substituting $\mu = 35$, $\mu^\prime = 100$, $\sigma = 10$, and $\sigma^\prime = 15$,

\begin{align*}
y(x)\ &= 1.5x + 47.5\ \rightarrow\ \{a = 47.5,\ b = 1.5\} &\ \textbf{(a)}
\end{align*}

To find the range, we need to insert the limits of $x$, 0 and 50:

\begin{align*}
y(0)\ &=\ 1.5(0) + 47.5\ \rightarrow y_{\mathrm{min}}=47.5 \\
y(50)\ &=\ 1.5(50) + 47.5\ \rightarrow y_{\mathrm{max}}=122.5 \\
\end{align*}

Thus, the range of $y$ is $\left[47.5, 122.5\right]$. **(b)**

A plot of the linear transformation follows: **(c)**

```{r question7, echo=FALSE, fig.height=4}
```

# Question 8

# Question 9
