---
title: "Homework 4"
subtitle: "IQM"
date: "`r strftime(Sys.Date(), format = '%Y-%m-%d')`"
author: Patrick Anker
header-includes:
output: 
  bookdown::pdf_document2:
    includes:
      in_header: "../preamble.tex"
    toc: false
    extra_dependencies: "float"
---

```{r setup1, include=FALSE, cache=FALSE}
knitr::read_chunk(here::here("homework4/analysis.R"))
```

```{r preamble, include=FALSE}
```

# Question 1 {-}

```{r question1, echo=FALSE}
```

```{r q1f1, echo=FALSE, out.width="55%", fig.align="center", fig.cap="Examination of maneuver performance (a)"}
```

Fig. \@ref(fig:q1f1) shows the comparison of scores, and they follow a linear trend. This is to be expected since the scores are both derived from a natural ability plus some noise ($\sigma=1.5$).

```{r q1t1, include=FALSE}
```

```{r q1t1-out, echo=FALSE}
kableExtra::kable(
  tab_q1t1,
  booktabs = TRUE,
  escape = FALSE,
  caption = "Extreme values are regressed to the mean"
) |>
  kableExtra::kable_styling(position = "center")
```

However, Tab. \@ref(tab:q1t1-out) shows that the mean score differences are not approximately zero (save for the "None" case). The negative feedback cases show a positive mean score difference, whereas the positive feedback cases show a negative mean score difference. This is purely a numerical artifact of averages: averages create shrinkage from extreme values. However, they are only *point estimates*. The full picture must include the reality that each score pair is drawn from a distribution, which has inherent noise.

# Question 2 {-}

```{r question2-a, echo=FALSE}
```

```{r q2t1, echo=FALSE}
kableExtra::kable(
  tab_q2t1,
  booktabs = TRUE,
  caption = "Comparison of methods for calculating mean bass fish length"
) |>
  kableExtra::kable_styling(position = "center")
```

```{r question2-b, echo=FALSE}
```

```{r q2t2, echo=FALSE}
kableExtra::kable(
  tab_q2t2,
  booktabs = TRUE,
  caption = "Comparison of methods for calculating difference of mean incomes by age group"
) |>
  kableExtra::kable_styling(position = "center")
```

